Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-23 TACAS2018/paper.tex
edited the synthesis sec

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Using optimization refinement for CEGIS of control systems}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In this section, we describe our technique for synthesizing safe
digital feedback controllers using CEGIS. Broadly speaking, CEGIS
denotes an iterative technique where each iteration performs inductive
generalisation based on counterexamples provided by a verification
engine.

We build upon this paradigm to create CEGORS (Counter Example Guided, Optimization
Refined Synthesis).  The main advantage of this approach, is that by optimizing the
synthesis results with respect to some objective function, we are able to reduce the
number of times the loop iterations, thus improving its performance.  Furthermore,
because later stages of synthesis will use more complex abstractions, the reduction
is on increasingly expensive formulas and therefore the improvent can be exponential.
%% In order to synthesize closed-loop digital control systems, we use a program
%% synthesis engine.  Our program implements Counter-Example Guided
%% Inductive Synthesis (CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}.  We
%% start by presenting its general architecture followed by describing the
%% parts specific to closed-loop control systems.

@@

Our approach is illustrated in Figure~\ref{fig:CEGORS}. It relies on the
construction of an unsound abstract model ({\sc ABSTRACT} phase) that approximates the dynamics
of the cyber-physical system in order to synthesise a candidate controller
and observer and then verify their behaviour against a sound over-approximation
of the concrete dynamics of the system for an infinite space over an infinite
time horizon.
The verifier's abstraction uses a CEGAR loop in order to minimise 
over-approximation errors and provides counterexamples for both
discrete and continuous dynamics to feed a CEGIS loop on the synthesiser
waiting for a new candidate solution.

Our CEGORS loop is similar to CEGIS with solution generalization~\cite{david2017program}.
The novelty of our approach relies on the synthesis refinement phase, which
enables us to generalise an existing candidate by means of optimisation. Rather
than allow unsuitable candidates reach the verification phase, and possibly
require further induction on the synthesiser, we use an optimizing algorithm
which finds a local minima for the properties of interest.  This reduces the
chances of the optimised candidate failing verification.  Furthermore, it
allows us to relax the quality of the synthesis, since unlikely candidates will
still be optimised during this phase. 
The optimization is performed on an objective function which will typically depend
on a subset of the properties.  Hence the chosen optimization function depends
on which properties fail verification. In our case, the minimal abstraction
used by the synthesiser almost always ensures stability, so our objective function
focuses on safety, although we manage to do optimizations that
simultaneously target safety, stability and response time. This is done by
making the dynamics as convergent as possible while maintaining the
verification properties that exclude excesively convergent ones.

@@

The CEGORS loop for this approach is described next.

@@

\begin{enumerate}
\item 
The {\sc ABSTRACT} module starts by creating an abstraction from the initial specification,
\begin{enumerate}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:FWL_effects}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{enumerate}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.

\item In {\sc synthesize}, we synthesize a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
\item {\sc Verify Synthesis} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy the tighter
specification, it feeds back to the {\sc Optimise} phase which optimises
an objective functiion in order to minimise the magnitude of the error in the verification.
If the candidate is valid, we feed it to {\sc Discrete Time}.

\item {\sc Optimise} 
% starts by taking $\mat{K},\mat{L} \not \models \phi_{static}$
% and synthesises $\mat{K}_{\langle C \rangle} = \mat{K}+\mat{\delta}_K, \mat{L}_{\langle C \rangle}=\mat{L}+\mat{\delta}_L$,
% where $\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \models \phi_{static}$.  
% If $\mat{K},\mat{L} \models \phi_{static}$ but we have a counterexample for $\mat{K},\mat{L} \models \phi$,
% then we proceed to optimise the controller.

Our optimization is iterative, following two stages.  The first stage seeks to
make the dynamics as convergent as possible, since convergent dynamics are
less likely to stray outside of the safe region (we say less likely because the
change in the eigenvectors changes the shape of the space and may cause some 
trajectories to go further).  In this case, we take the largest eigenvalue and
reduce it by a factor (smaller than .5) that still satisfies $\phi_{static}$.  The
resulting observer controller is then verified against the full specification and
the maximum deviation from the safe region is measured (maximum distance from
the discovered reach set to the safety region, measured perpendicularly to any
of the faces of the safe region). 
While this number is monotonically decreasing we continue to iterate until
it reaches a negative value (i.e. safe). If the number increases, we change
our objective function to the second stage.

In the second stage, at each iteration we take the complex eigenvalue
with the largest phase (largest angle in the complex space), and reduce its
imaginary component while maintaining the specification $\phi_{static}$.
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and may have a
smaller rotation, which facilitates meeting the safety specification (we note
that real convergent eigenvalues are trivially safe).

If we cannot synthesise a suitable candidate using optimisation, we signal the
{\sc Abstract} phase to refine $\phi_{synth}$ and request {\sc Synthesise} to
produce a new candidate.  Otherwise the refined candidate solution is given back
to {\sc Verify Synthesis}.
%
\item {\sc Discrete Time} verifies $\phi_{safety}^{\infty}$ (unbounded time)
using discrete time dynamics.  If the verification fails, we report the
counterexample to {\sc Refine Abstraction} including details of initial states
and iterations that failed the specification.  
\item {\sc Refine Abstraction} changes $\phi_{safety}$ to encompass any arising
counterexample.  This can be done by increasing the time horizon $k$ for the
safety specification of the violating initial condition and input.
It also modifies $\phi_{static}$ to account for the counterexample.
If the magnitude of the violation is small, {\sc Optimise} may find a new
candidate, otherwise it will feed through to {\sc Abstract}.
%
\item {\sc Continuous Time} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
\item {\sc Sample} will select a smaller suitable sample time and modify the abstractions to meet the discretization. 
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%

\item If we reach the {\sc abstract} phase, it means that the candidate
  solution is not valid, in which case we must refine $\phi_{synth}$. If $\phi_{synth}$ cannot be further refined, 
  then we cannot synthesise a new candidate and return a synthesis failure.
\end{enumerate}

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\subsection{Using optimization refinement for CEGIS of control systems}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In this section, we describe our technique for synthesizing safe
digital feedback controllers using CEGIS. Broadly speaking, CEGIS
denotes an iterative technique where each iteration performs inductive
generalisation based on counterexamples provided by a verification
engine.

We build upon this paradigm to create CEGORS (Counter Example Guided, Optimization
Refined Synthesis) \note{not sure whether we need an acronym for this, we could just call it
CEGIS with optimization and abstraction refinement}.
Our approach is illustrated in Figure~\ref{fig:CEGORS} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer \note{are these obeying the stability spec?}.
\item {\sc model}
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

The verifier's abstraction uses a CEGAR loop in order to minimise 
over-approximation errors and provides counterexamples for both
discrete and continuous dynamics to feed a CEGIS loop on the synthesiser
waiting for a new candidate solution. \note{not sure about this sentence.
  Which is the CEGAR loop?}

The main advantage of this approach is two-fold:
\begin{itemize}
\item[1.] Optimizing the candidate solution in each
  iteration of the CEGIS loop by computing the local optimum
  enables us to obtain high quality counterexamples in the {\sc verify} phase.
  In turn, this has the potential to reduce the number of iteration of the CEGIS
  loop required to obtain a solution.
  \note{probably explain high quality counterexample}
\item[2.] We start with simpler (i.e. imprecise)
  abstractions that get
  refined into more complex ones as the procedure progresses. Thus, solutions
  that don't require a high level of precision are obtained fast. 
\end{itemize}

%% because later stages of synthesis will use more complex abstractions, the reduction
%% is on increasingly expensive formulas and therefore the improvent can be exponential.

%% In order to synthesize closed-loop digital control systems, we use a program
%% synthesis engine.  Our program implements Counter-Example Guided
%% Inductive Synthesis (CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}.  We
%% start by presenting its general architecture followed by describing the
%% parts specific to closed-loop control systems.

@@

%%  It relies on the
%% construction of an unsound abstract model ({\sc ABSTRACT} phase) that approximates the dynamics
%% of the cyber-physical system in order to synthesise a candidate controller
%% and observer and then verify their behaviour against a sound over-approximation
%% of the concrete dynamics of the system for an infinite space over an infinite
%% time horizon.

%Our CEGORS loop is similar to CEGIS with solution generalization~\cite{david2017program}.

%% The novelty of our approach relies on the synthesis refinement phase, which
%% enables us to generalise an existing candidate by means of optimisation.
%% Rather
%% than allow unsuitable candidates reach the verification phase, and possibly
%% require further induction on the synthesiser, we use an optimizing algorithm
%% which finds a local minima for the properties of interest.  This reduces the
%% chances of the optimised candidate failing verification.  Furthermore, it
%% allows us to relax the quality of the synthesis, since unlikely candidates will
%% still be optimised during this phase. 
%% The optimization is performed on an objective function which will typically depend
%% on a subset of the properties.  Hence the chosen optimization function depends
%% on which properties fail verification.

During the optimization, the minimal abstraction used by the synthesiser almost
always ensures stability, so our objective function focuses on safety,
although we manage to do optimizations that simultaneously target
safety, stability and response time.
\note{Are all the
  candidates computed by {\sc synthesize} stable? State clearly if
  so. You say ``almost always''.}
This is done by making the
dynamics as convergent as possible while maintaining the verification
properties that exclude excesively convergent ones.

@@

@@

%\begin{enumerate}
%\item

The CEGORS loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification} \note{broken reference}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:FWL_effects} \note{broken reference}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

%\item
\paragraph{4. {\sc Optimise}} computes the local optimum in the neighborhood of
the current candidate solution. \note{is this really a local optimum?}
% starts by taking $\mat{K},\mat{L} \not \models \phi_{static}$
% and synthesises $\mat{K}_{\langle C \rangle} = \mat{K}+\mat{\delta}_K, \mat{L}_{\langle C \rangle}=\mat{L}+\mat{\delta}_L$,
% where $\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \models \phi_{static}$.  
% If $\mat{K},\mat{L} \models \phi_{static}$ but we have a counterexample for $\mat{K},\mat{L} \models \phi$,
% then we proceed to optimise the controller.
%
Our optimization is iterative, following two stages.  The first stage seeks to
make the dynamics as convergent as possible, since convergent dynamics are
less likely to stray outside of the safe region (we say less likely because the
change in the eigenvectors changes the shape of the space and may cause some 
trajectories to go further).
\note{Can we formally define the objective function?}

In this case, we take the largest eigenvalue and reduce it by a factor
(smaller than .5) that still satisfies $\phi_{static}$.  The resulting
observer controller is then verified against the full specification
\note{full specification?}
and the maximum deviation from the safe region is measured (i.e. maximum
distance from the discovered reach set to the safety region, measured
perpendicularly to any of the faces of the safe region).  While this
number is monotonically decreasing we continue to iterate until it
reaches a negative value (i.e. safe). If the number increases, we
change our objective function to the second stage.
\note{again, formal description? what is the objective function in this stage?}

In the second stage, at each iteration we take the complex eigenvalue
with the largest phase (largest angle in the complex space), and reduce its
imaginary component while maintaining the specification $\phi_{static}$.
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and may have a
smaller rotation, which facilitates meeting the safety specification
\note{$\phi_{safety}$?}
(we note
that real convergent eigenvalues are trivially safe).

If we cannot optimize the candidate such that it passes the {\sc
  verify synthesis} check, we signal the {\sc Abstract} phase to
refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new
candidate.  Otherwise the refined candidate solution is given back to
{\sc Verify Synthesis}.
%
%\item
\paragraph{5. {\sc Discrete Time}} verifies $\phi_{safety}^{\infty}$ (unbounded time)
using discrete time dynamics.  If the verification fails, we report the
counterexample to {\sc Refine Abstraction} including details of initial states
and iterations that failed the specification.  
%\item
\paragraph{6. {\sc Refine Abstraction}} changes $\phi_{safety}$ to encompass any arising
counterexample.  This can be done by increasing the time horizon $k$ for the
safety specification of the violating initial condition and input.
It also modifies $\phi_{static}$ to account for the counterexample.
If the magnitude of the violation is small, {\sc Optimise} may find a new
candidate, otherwise it will feed through to {\sc Abstract}.
%
%\item
\paragraph{7. {\sc Continuous Time}} verifies that the
transitions between sample times meet the specification $\phi_{safety}^{\infty}$.  If
a counterexample is found, information about the maximum distance of the
violation to the specification boundary is fed to {\sc Sample}. 
%\item
\paragraph{8. {\sc Sample}} selects a smaller suitable sample time and modify the abstractions to meet the discretization. 
It will then feed back to the abstraction refinement loop to look for a suitable action for the updated model.
%

%\item
If we reach the {\sc abstract} phase, it means that the candidate
  solution is not valid, in which case we must refine $\phi_{synth}$. If $\phi_{synth}$ cannot be further refined, 
  then we cannot synthesise a new candidate and return a synthesis failure.
%\end{enumerate}
